---
title: "Data Mining Project"
author: "Fatemeh Golpour"
date: "2024-04-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries
```{r Libraries}
library(ggplot2)
library(reshape2)
library(dplyr)
library(tidyr)
library(tidytext)
library(quanteda)
library(caret)
library(arules)
library(arulesViz)
library(randomForest)
```


## Data Understanding

```{r}
anime_data <- read.csv("Anime.csv")
summary(anime_data)
```

#```{r}
#head(anime_data)

#```

```{r}
dim(anime_data)
```

```{r}
# Find numerical features
numerical_features <- sapply(anime_data, is.numeric)
# Filter numerical data
numerical_data <- anime_data[, numerical_features]
```
Plot histograms of the numerical data:

```{r}
# Plotting Layout: 3 rows and 3 column
par(mfrow = c(3, 3))
for (col in colnames(numerical_data)) {
    hist(numerical_data[[col]],
         main = paste("Histogram of", col),  
         xlab = col,  
         ylab = "Frequency",  
         col = "lightblue", 
         breaks = 20)
}
```
Plot correlation heatmap of the numerical data:


```{r}
correlation_matrix <- cor(numerical_data, use = "pairwise.complete.obs")

# Reshape correlation matrix
melted_corr_matrix <- melt(correlation_matrix)

heatmap <- ggplot(melted_corr_matrix, aes(x = Var1, y = Var2, fill = value)) +
           geom_tile() +
           scale_fill_gradient(low = "blue", high = "black") +
           labs(title = "Correlation Heatmap of Numerical Data") +
           theme_minimal()+
           # Rotate the x-axis text (column names) to be vertical
           theme(axis.text.x = element_text(angle = 90, hjust = 1)) 

print(heatmap)
```
## Data Preparation

Check for missing values in the dataset:

```{r}

missing_per_column <- colSums(is.na(anime_data))

print("Missing values per column:")
print(missing_per_column)

```
```{r}
missing_episodes <- anime_data[is.na(anime_data$Episodes), ]

```

Replace the the missing values in Episodes with 1, if the column 'Type' is equal to: OVA, Movie or Special, otherwise replace them with the average of the column:
```{r}
# Calculate the average of Episodes for non-OVA, Movie, or Special types
average_episodes <- mean(anime_data$Episodes[!anime_data$Type %in% c("OVA", "Movie", "Special")], na.rm = TRUE)

# Replace missing values in Episodes
anime_data <- anime_data %>%
  mutate(Episodes = case_when(
    is.na(Episodes) & Type %in% c("OVA", "Movie", "Special") ~ as.integer(1),
    is.na(Episodes) ~ as.integer(average_episodes),
    TRUE ~ Episodes
  ))
```

Replace the missing values in Score, Scored_Users, Ranked and Duration_Minutes with the average of the columns.

```{r}
# Calculate the average of each column
average_score <- mean(anime_data$Score, na.rm = TRUE)
average_scored_users <- mean(anime_data$Scored_Users, na.rm = TRUE)
average_ranked <- mean(anime_data$Ranked, na.rm = TRUE)
average_duration_minutes <- mean(anime_data$Duration_Minutes, na.rm = TRUE)


# Replace missing values in Score, Scored_Users, and Ranked with the average of each column
anime_data <- anime_data %>%
  mutate(
    Score = if_else(is.na(Score), average_score, Score),
    Scored_Users = if_else(is.na(Scored_Users), average_scored_users, as.numeric(Scored_Users)),
    Ranked = if_else(is.na(Ranked), average_ranked, as.numeric(Ranked)),
    Duration_Minutes = if_else(is.na(Duration_Minutes), average_duration_minutes, as.numeric(Duration_Minutes))

  )
```


Identify 'Unknown' values:

```{r}
unknown_counts <- sapply(anime_data, function(column) sum(column == "Unknown"))

#print("Number of 'unknown' values per column:")
#print(unknown_counts)

```

Since generating random text might affect the accuracy of our models, we do not change the unknown values.

```{r}
str(anime_data)

```

```{r}
text_columns <- colnames(anime_data)[sapply(anime_data, is.character)]

```
Clean the text columns:
```{r}
clean_text_columns <- function(data, text_columns) {
    for (col in text_columns) {
        # Tokenize the data and clean it
        tokens_col <- tokens(data[[col]], remove_punct = TRUE, remove_symbols = TRUE) %>%
            tokens_tolower() %>%
            tokens_remove(pattern = "[^a-zA-Z\\s]") %>%
            tokens_remove(pattern = stopwords("en"))
        
        # Convert tokens back to single strings
        data[[col]] <- sapply(tokens_col, paste, collapse = " ")
        
        # Handle missing values
        data[[col]][is.na(data[[col]])] <- "Unknown"
    }
    
    return(data)
}

text_columns <- c("Title", "Synonyms", "Japanese", "English", "Synopsis", "Start_Aired", "End_Aired", "Premiered", "Broadcast", "Producers", "Licensors", "Studios", "Genres", "Themes", "Demographics" )

anime_data_cleaned_text <- clean_text_columns(anime_data, text_columns)

```


Convert the potential categorical columns to factors:

```{r}
categorical_columns <- c("Type", "Status", "Source", "Rating")

for (col in categorical_columns) {
        anime_data_cleaned_text[[col]] <- as.factor(anime_data_cleaned_text[[col]])
}
```

Plot the frequencies of the categorical columns:

```{r}
categorical_columns <- c("Type", "Status", "Source", "Rating")

for (column in categorical_columns) {
    plot <- ggplot(anime_data_cleaned_text, aes_string(x = column)) +
        geom_bar(fill = "blue") +  
        labs(title = paste("Distribution of", column),
             x = column,
             y = "Frequency") +
        theme(axis.text.x = element_text(angle = 90, hjust = 1))
    
    print(plot)
}

```

We can observe that our data is imbalanced in our different categorical variables.
Seems like performing a classification task to determine the rating of an anime could be interesting.

```{r}
ggplot(anime_data_cleaned_text, aes(x = "", fill = Rating)) +
    geom_bar(width = 1, stat = "count") +
    coord_polar(theta = "y") +
    labs(title = paste("Distribution of ratings")) +
    theme_void()

```

## Modelling

  Create a new dataframe with the more relevant features to work with:

```{r}
anime_df <- data.frame(English=anime_data_cleaned_text$English, Synopsis=anime_data_cleaned_text$Synopsis,
                       Studios=anime_data_cleaned_text$Studios, Genres=anime_data_cleaned_text$Genres,
                       Themes=anime_data_cleaned_text$Themes, Demographics=anime_data_cleaned_text$Demographics,
                       Type=anime_data_cleaned_text$Type, Status=anime_data_cleaned_text$Status,
                       Source=anime_data_cleaned_text$Source, Score=anime_data$Score, Scored_Users=anime_data$Scored_Users,
                       Ranked=anime_data$Ranked, Rating=anime_data_cleaned_text$Rating
                       )
# Columns you want to convert from factor to character
columns_to_convert <- c("Type", "Status", "Source")

# Convert specified factor columns to character
anime_df[columns_to_convert] <- lapply(anime_df[columns_to_convert], as.character)
```

```{r}
new_text_columns <- c("Type", "Status", "Source")

anime_df <- clean_text_columns(anime_df, new_text_columns)

```

### Association Rule Mining
To inspect the association rules, in order to demonstrate the rules better, I decided to use the three columns that have repetitive words in them but with different combinations: Genres, Themes, and Demographics.
```{r}
data_categorical <- data.frame(anime_df$Genres, anime_df$Themes, anime_df$Demographics)

df_filtered <- data_categorical %>%
  filter_all(all_vars(. != "unknown"))

# Convert categorical columns to factors 
categorical_cols <- sapply(df_filtered, is.character)
df_filtered[categorical_cols] <- lapply(df_filtered[categorical_cols], as.factor)


# Convert the data frame to a transaction object
transactions <- as(df_filtered, "transactions")

# Perform association rule mining using the apriori algorithm
rules <- apriori(transactions, parameter = list(support = 0.01, confidence = 0.5))

# View the association rules
inspect(rules)

```

Plot the association rules:

```{r}
# Scatter plot of support vs confidence
plot(rules, measure = c("support", "confidence"), shading = "lift", col = c("red", "blue"))
```
```{r}
# Parallel coordinate plot
plot(rules, method = "paracoord", control = list(reorder = TRUE), col = c("red", "blue", "green"))
```

### Classification Task - Random Forest

```{r}

text_columns <- c("English", "Genres", "Themes", "Demographics", "Type", "Status", "Source")
# Convert textual data into a DTM(document-term matrix)
dtm <- lapply(anime_df[text_columns], function(col) {
    tokens_col <- tokens(col)
    dfm_col <- dfm(tokens_col)
    return(dfm_col)
})
combined_dfm <- Reduce(cbind, dtm)
# Combine the DTM with other features 
features <- convert(combined_dfm, to = "data.frame")
# Prepare the labels for classification
features$Rating <- as.factor(anime_df$Rating)

duplicate_columns <- colnames(features)[duplicated(colnames(features))]
print(duplicate_columns)
if (length(duplicate_columns) > 0) {
    for (i in seq_along(duplicate_columns)) {
        col_index <- which(colnames(features) == duplicate_columns[i])
        for (j in seq_along(col_index)) {
            colnames(features)[col_index[j]] <- paste0(duplicate_columns[i], "_", j)
        }
    }
}

trainIndex <- createDataPartition(features$Rating, p = .7, list = FALSE)
trainData <- features[trainIndex, ]
testData <- features[-trainIndex, ]

# Train a classification model - Random Forest
model <- train(Rating ~ ., data = trainData, method = "rf",
               tuneGrid = expand.grid(mtry = c(2, 3, 4)),
               trControl = trainControl(method = "cv", number = 5))

# Evaluate the model
predictions <- predict(model, testData)
confusionMatrix(predictions, testData$Rating)

```

Unfortunately, I don't have enough computing power to run this on the complete textual data(synopsis) and get the results I want, as my laptop crashed several times during my attempts.
After this, I was planning to use the textual data to perform a regression task and predict the numerical data(Score, Scored_Users, and Ranked).
